Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-01-30 Report/decomposition_solutions.tex
cooked half only

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. Let us consider the \review{convex subproblem with the fixed receive beamformers \me{\mvec{w}{l,k,n}} presented} in \eqref{eqn-9} based on the Taylor series approximation for the nonconvex constraint. The following discussions are equally valid for the \ac{MSE} based solution outlined in \eqref{eqn-mse-2} as well. Since the objective of \eqref{eqn-9} can be decoupled across each \ac{BS}, the centralized problem can be equivalently written as
{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e}, \IEEEyessubnumber
\end{IEEEeqnarray}}
where \me{\tilde{\mbf{v}}_b} denotes the vector of of weighted queue deviation corresponding to users \me{k \in \mc{U}_b}.

Following the approaches presented in \cite{pennanen2011decentralized,tolli2011decentralized}, the coupling constraint \eqref{eqn-9.1c} or \eqref{eqn-mse-1.3} can be expressed by grouping the interference contribution from each \ac{BS} in \me{\mathcal{B}} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl}\label{eqn-decent-3}
\beta_{l,k,n} & \geq & \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&\quad& + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b} \; + \; \enoise,
\end{IEEEeqnarray}}{\allowdisplaybreaks
\begin{multline}\label{eqn-decent-3}
	\enoise + \sum_{\substack{j = 1,j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b}  \\
	+ \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 \leq \beta_{l,k,n},
\end{multline}}
where \me{\zeta_{l,k,n,b}} is the total interference caused by the transmission of \ac{BS} \me{b} to user \me{k \in \mc{U}_{b_k}} in the spatial stream \me{l} and sub-channel \me{n}. It is given by the following upper bound as
\begin{equation}
\zeta_{l,k,n,b} \geq \sum_{i \in \mc{U}_b} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2, \forall b \in \bar{\mc{B}}_{b_k}.
\end{equation}

The decentralization is achieved by decomposing the original convex problem in \eqref{eqn-decent-1} by an iterative master-slave model using either primal or \acl{DD} method. The coupling variables are controlled by the master problem and the slave problems solve for the optimization variables associated with itself only. Now, to decouple the problem in \eqref{eqn-decent-1}, the coupling variable \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3}, which bind all \acp{BS} in \me{\mc{B}}, is replaced by a \ac{BS} specific variable \me{\zeta_{l,k,n,b}^{\{b_k\}}}. Using the \ac{BS} specific variable \me{\zeta_{l,k,n,b}^{\{b_k\}}} in \eqref{eqn-decent-3}, the subproblem in \eqref{eqn-decent-1} decouples completely for each \ac{BS} to find the precoders only for the users associated with the \ac{BS} \me{b_k}. The slave subproblems with the \ac{BS} specific interference variables are controlled by a master subproblem, which determines the interference variable \me{\zeta_{l,k,n,b}^{\{b_k\}}} in accordance with the global objective.

>>>>

Let us consider the \review{convex subproblem with the fixed receive beamformers \me{\mvec{w}{l,k,n}} presented} in \eqref{eqn-9} based on the Taylor series approximation for the nonconvex constraint. The following discussions are equally valid for the \ac{MSE} based solution outlined in \eqref{eqn-mse-2} as well. Since the objective of \eqref{eqn-9} can be decoupled across each \ac{BS}, the centralized problem can be equivalently written as
{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e}, \IEEEyessubnumber
\end{IEEEeqnarray}}
where \me{\tilde{\mbf{v}}_b} denotes the vector of of weighted queue deviation corresponding to users \me{k \in \mc{U}_b}.

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. Following the approaches presented in \cite{pennanen2011decentralized,tolli2011decentralized}, the coupling constraint \eqref{eqn-9.1c} or \eqref{eqn-mse-1.3} can be expressed by grouping the interference contribution from each \ac{BS} in \me{\mathcal{B}} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl}\label{eqn-decent-3}
\beta_{l,k,n} & \geq & \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&\quad& + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b} \; + \; \enoise,
\end{IEEEeqnarray}}{\allowdisplaybreaks
\begin{multline}\label{eqn-decent-3}
	\enoise + \sum_{\substack{j = 1,j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,k,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta_{l,k,n,b}  \\
	+ \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{j,i,n} |^2 \leq \beta_{l,k,n},
\end{multline}}
where \me{\zeta_{l,k,n,b}} is the total interference caused by the transmission of \ac{BS} \me{b} to user \me{k \in \mc{U}_{b_k}} in the spatial stream \me{l} and sub-channel \me{n}. It is given by the following upper bound as
\begin{equation}
\zeta_{l,k,n,b} \geq \sum_{i \in \mc{U}_b} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2, \forall b \in \bar{\mc{B}}_{b_k}.
\end{equation}

The decentralization is achieved by decomposing the original convex problem in \eqref{eqn-decent-1} by a parallel iterative subproblems using either primal or \acl{DD} method. The coupling variables are updated in each iteration by exchanging limited information among the subproblems. Now, to decouple the problem in \eqref{eqn-decent-1}, the coupling variable \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3}, which bind all \acp{BS} in \me{\mc{B}}, is replaced by a \ac{BS} specific variable \me{\zeta_{l,k,n,b}^{\{b_k\}}}. Using the \ac{BS} specific variable \me{\zeta_{l,k,n,b}^{\{b_k\}}} in \eqref{eqn-decent-3}, the subproblem in \eqref{eqn-decent-1} decouples completely for each \ac{BS} to find the precoders for the associated users by exchanging limited information to update \me{\zeta_{l,k,n,b}^{\{b_k\}}} in each iteration for the \ac{BS} \me{b_k}.

%The subproblems with the \ac{BS} specific interference variables are controlled by a master subproblem, which determines the interference variable \me{\zeta_{l,k,n,b}^{\{b_k\}}} in accordance with the global objective.

### 2015-01-30 Report/primal_approach.tex
cooked half only

The \acl{PD} approach decomposes the problem by fixing the interference variables \me{\zeta_{l,k,n,b} \forall k,b} in order to perform the precoder design independently across each \ac{BS}. Once the optimal precoders are designed at each \ac{BS} with the fixed interference constraints \eqref{eqn-decent-3}, the dual variables corresponding to the interference constraints are exchanged between the cooperating \acp{BS} in \me{\mc{B}} to update the interference variables \me{\zeta_{l,k,n,b}} for the next iteration until convergence. The primal approach is discussed extensively for the min-power problem in \cite{pennanen2011decentralized} and much of the current work follows similar approach. 
%Much of the details are provided in the Appendix \ref{a-2}.

\subsubsection*{Convergence}
The convergence of the primal decomposition is similar to that of the centralized problem if the interference variables \me{\zeta_{l,k,n,b}} are allowed to converge to a stationary point. In practice, we can limit the number of exchanges to \me{J_{\max}} after which the \ac{SCA} update is performed until convergence or for \me{I_{\max}} times. The update of \me{\tilde{p}_{l,k,n}, \tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} can be made in conjunction with the receiver update \me{\mvec{W}{k,n}}. The receiver update can be made by using the precoded pilot transmission from each user as in \cite{komulainen2013effective}.

>>>>

In \acl{PD}, the convex problem in \eqref{eqn-decent-1} is solved for the optimal transmit precoders in an iterative manner for a fixed \ac{BS} specific interference terms \me{\zeta_{l,k,n,b}^{\{b_k\}}} using master-slave model \cite{pennanen2011decentralized}. The slave subproblem is solved in each \ac{BS} for the optimal transmit precoders for the associated users only by assuming a fixed interference \me{\zeta_{l,k,n,b}^{\{b_k\}}} from other \acp{BS} \me{b \in \mc{B} \backslash b_k} in each iteration. Upon finding the optimal associated transmit precoders by each slave subproblems, the master problem is used to update the \ac{BS} specific interference terms \me{\zeta_{l,k,n,b}^{\{b_k\}}} to a new value by using dual variables corresponding to the interference constraint \eqref{eqn-decent-3} as discussed in \cite{pennanen2011decentralized}. In this manner, the interference variables are updated until the global consensus is obtained. The primal approach is similar to the minimum power precoder design presented in \cite{pennanen2011decentralized}. Note that the master problem considers \me{\zeta_{l,k,n,b}^{\{b_k\}} \forall \{b,b_k\} \in \mc{B}} as a variable and the slave subproblems assume this to be a constant in each iteration that will be updated by the master problem.

\begin{comment}
The \acl{PD} approach decomposes the problem by fixing the interference variables \me{\zeta_{l,k,n,b} \forall k,b} in order to perform the precoder design independently across each \ac{BS}. Once the optimal precoders are designed at each \ac{BS} with the fixed interference constraints \eqref{eqn-decent-3}, the dual variables corresponding to the interference constraints are exchanged between the cooperating \acp{BS} in \me{\mc{B}} to update the interference variables \me{\zeta_{l,k,n,b}} for the next iteration until convergence. The primal approach is discussed extensively for the min-power problem in \cite{pennanen2011decentralized} and much of the current work follows similar approach. 
%Much of the details are provided in the Appendix \ref{a-2}.

\subsubsection*{Convergence}
The convergence of the primal decomposition is similar to that of the centralized problem if the interference variables \me{\zeta_{l,k,n,b}} are allowed to converge to a stationary point. In practice, we can limit the number of exchanges to \me{J_{\max}} after which the \ac{SCA} update is performed until convergence or for \me{I_{\max}} times. The update of \me{\tilde{p}_{l,k,n}, \tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} can be made in conjunction with the receiver update \me{\mvec{W}{k,n}}. The receiver update can be made by using the precoded pilot transmission from each user as in \cite{komulainen2013effective}. 
\end{comment}
